\documentclass[12pt]{article}

\usepackage[margin=1.0in]{geometry}
\usepackage[parfill]{parskip}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage[font=bf]{caption}
\usepackage{changepage}
\usepackage{hyperref}



%-------------------------------------------------------------------------------
\begin{document}
%-------------------------------------------------------------------------------



\setlength{\parskip}{12pt}      % set paragraph skipping to one whole line
\setlength{\tabcolsep}{12pt}    % give some room between table columns
\hfuzz=60pt                     % give a little margin room for the double-column tables



% establish a condensed itemized list format
\newenvironment{myitemize}
{ \begin{itemize}
    \setlength{\itemsep}{0pt}
    \setlength{\parskip}{0pt}
    \setlength{\parsep}{0pt}     }
{ \end{itemize}                  } 



% establish theorem and lemma formats
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}



\begin{titlepage}
    \begin{center}
        \vspace*{\fill}
        
        \huge
        The Kalman Filter \\
        
        \large
        An Adventure in Derivation
        
        \vspace{24pt}
        
        \large
        Robert S. Huston \\
        Pinpoint Dynamics, LLC \\
        
        \vspace{24pt}
        
        \today
        
        \vspace{24pt}
        
        \large
        Copyright © 2024, 2025 Robert S. Huston. All rights reserved.
        
        \vfill
    \end{center}
\end{titlepage}



\begin{abstract}
This document presents an assorted collection of derivations and notes pertaining to the
Kalman filter. They are targeted to the practicing engineer who has been tasked with
developing and implementing a Kalman filtering solution. The mathematical derivations are
intended to eliminate the mysteries of how and why the Kalman filter works. In addition,
the document presents the various versions of the Kalman filter and offers some key
implementation strategies.
\end{abstract}



\clearpage



\begin{adjustwidth}{1in}{1in}
    \vspace*{\fill}
    
    \textit{
        So I jump ship in Hong Kong and I make my way over to Tibet, and I get on as a
        looper at a course over in the Himalayas.
        A looper, you know, a caddy, a looper, a jock. So, I tell them I'm a pro jock,
        and who do you think they give me? The Dalai Lama, himself. Twelfth son of the Lama.
        The flowing robes, the grace, bald...striking. So, I'm on the first tee with him.
        I give him the driver. He hauls off and whacks one - big hitter, the Lama - long,
        into a ten-thousand foot crevasse, right at the base of this glacier.
        Do you know what the Lama says? "Gunga galunga...gunga, gunga-lagunga."
        So we finish the eighteenth and he's gonna stiff me. And I say, "Hey, Lama, hey,
        how about a little something, you know, for the effort, you know."
        And he says, "Oh, uh, there won't be any money, but when you die, on your deathbed,
        you will receive total consciousness." So I got that goin' for me, which is nice.
    }
    
    - Carl Spackler (Bill Murray, "Caddyshack" - 1980)
    
    \vfill
\end{adjustwidth}



\clearpage



\tableofcontents



\clearpage



\include{introduction}
\clearpage

\include{conventions-and-notations}
\clearpage

\include{weighted-least-squares}
\clearpage

\include{kalman-filter}
\clearpage

\include{information-kalman-filter}
\clearpage

\include{steady-state-kalman-filter}
\clearpage

\include{relationship-to-least-squares}
\clearpage

\include{quadratic-factors}
\clearpage

\include{uncorrelated-observations}
\clearpage

\include{ud-factorized-kalman-filter}
\clearpage

\include{extended-kalman-filter}
\clearpage

\include{unscented-kalman-filter}
\clearpage

\include{references}
\clearpage



%-------------------------------------------------------------------------------
\end{document}
%-------------------------------------------------------------------------------
