\section{References}

This list comprises what I consider to be the essential references for my taming of
the Kalman filter. They are not all explicitly cited, but they do form a comprehensive
list of the background resources used for this document.

% Hide default bibliography References title
\renewcommand{\refname}{}
\vspace{-11mm}

\begin{thebibliography}{9}

\bibitem{kalman1960}
Kalman, R. E., 1960, \\
\emph{A New Approach to Linear Filtering and Prediction Problems}, \\
Trans. ASME, Journal of Basic Engineering, vol. 82, Series D, pp. 34-45, Mar. 1960

\bibitem{thornton1976}
Thornton, C. L., 1976, \\
\emph{Triangular Covariance Factorizations for Kalman Filtering}, \\
National Aeronautics and Space Administration, Technical Memorandum 33-798, Oct. 1976

\bibitem{bierman1977}
Bierman, G. J., 1977, \\
\emph{Factorization Methods for Discrete Sequential Estimation}, \\
Academic Press, Inc., New York, NY

\bibitem{rgbrown1983}
Brown, R. G., 1983, \\
\emph{Introduction to Random Signal Analysis and Kalman Filtering}, \\
John Wiley \& Sons, Inc., New York, NY.

\bibitem{sorenson1985}
Sorenson, H. W. (Ed.), 1985, \\
\emph{Kalman Filtering: Theory and Application}, \\
IEEE Press, New York, NY.

\bibitem{chuichen1987}
Chui, C. K. and Chen, G., 1987, \\
\emph{Kalman Filtering with Real-Time Applications}, \\
Springer-Verlag, New York, NY.

\bibitem{grewalandrews1993}
Grewal, M. S., and Andrews, A. P., 1993, \\
\emph{Kalman Filtering: Theory and Practice}, \\
Prentice-Hall, Englewood Cliffs, NJ.

\bibitem{julieruhlmann1997}
Julier, S. J., and Uhlmann, J. K., 1997, \\
\emph{A New Extension of the Kalman Filter to Nonlinear Systems}, \\
Proc. SPIE 3068, Signal Processing, Sensor Fusion, and Target Recognition VI, 28 July 1997; https://doi.org/10.1117/12.280797

\bibitem{brookes2020}
Brookes, M., 2020, \\
\emph{The Matrix Reference Manual}, \\
http://www.ee.imperial.ac.uk/hp/staff/dmb/matrix/intro.html

\end{thebibliography}
